---
title: "Karite genetic diversity"
author: "Karen Cristine Goncalves dos Santos"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
---

```{r Load packages, message=FALSE, warning=FALSE, include=FALSE}
packages_to_use = c(
	"poppr", "adegenet", "ape",
	"mmod", "treemap", "tidyverse",
	"vcfR", "pegas", "hierfstat",
	"ggpubr")

# install.packages(packages_to_use, repos = "http://cran.rstudio.com", 
# 								 dependencies = TRUE)
for (i in packages_to_use) { 
	library(i, character.only = T)}
load("RData/differentiation_metrics.RData")
load("RData/Genetic_parameters.RData")
load("RData/missing_data.RData")
```
```{r Missing data, echo=FALSE, fig.height=7, fig.width=6, message=FALSE, warning=FALSE}
theme_set(theme_bw())

# data.frame(locus = names(miss_loc), 
# 					 values = unname(miss_loc))  %>%
# 	ggplot(aes(y = locus, x = values)) +
# 	geom_col(fill = "black") + 
# 	labs(x = "Proportion of available data",
# 			 y = "Loci",
# 			 title = "Loci with missing data") +
# 	theme(axis.text.y = element_blank(),
# 				axis.ticks.y = element_blank())

# cat("    \n\n")

data.frame(locus = names(miss_80_loc), 
					 values = unname(miss_80_loc))  %>%
	ggplot(aes(y = locus, x = values)) +
	geom_col(fill = "black") + 
	labs(x = "Proportion of available data",
			 y = "Loci",
			 title = "Loci with > 20% of missing data") +
	theme(axis.ticks.y = element_blank())

cat("    \n\n")

# data.frame(individuals = names(miss_ind), 
# 					 values = unname(miss_ind))  %>%
# 	ggplot(aes(y = individuals, x = values)) +
# 	geom_col(fill = "black") + 
# 	labs(x = "Proportion of available data",
# 			 y = "Individuals",
# 			 title = "Individuals with missing data") +
# 	theme(axis.text.y = element_blank(),
# 				axis.ticks.y = element_blank())
# 
# cat("    \n\n")

data.frame(individuals = names(miss_80_ind), 
					 values = unname(miss_80_ind))  %>%
	ggplot(aes(y = individuals, x = values)) +
	geom_col(fill = "black") + 
	labs(x = "Proportion of available data",
			 y = "Individuals",
			 title = "Individuals with > 20% of missing data") +
	theme(axis.ticks.y = element_blank())

```


```{r message=FALSE, warning=FALSE, include=FALSE}
library(knitr)

new_basicstats <- list(
	countries = 
		kable(basic_stats_print[
			, c("acronym", "variable", groups[["countries"]])], 
					format = "simple",
					col.names = c("", "", groups[["countries"]])),
	use = basic_stats_print[, groups[["use"]]] %>% t() %>% 
		as.data.frame(row.names = groups[["use"]]) %>%
		kable(row.names = T,
					format = "simple",
					col.names = basic_stats_print$acronym)
	
)


for (i in groups$countries) {
	regions = groups$regions[groups$regions %in% 
													 	population_information$Region[
													 		population_information$Country == i]]
	new_basicstats[[i]][["regions"]] = 
		basic_stats_print[, regions] %>% t() %>% 
		as.data.frame(row.names = regions) %>%
		kable(row.names = T,
					format = "simple",
					col.names = basic_stats_print$acronym)
	
	if (i == "MALI") {next}
	pops = groups$populations[groups$populations %in% 
															population_information$Location.Population[
																population_information$Country == i]]
	
	new_basicstats[[i]][["pops"]] = 
		basic_stats_print[basic_stats_print$acronym %in% 
					 	c("Ho", "Hs", "Ht", "Fis"), pops] %>% 
		t() %>% as.data.frame(row.names = pops) %>%
		kable(row.names = T,
					format = "simple",
					col.names = c("Ho", "Hs", "Ht", "Fis"))
	
	
}
```

## Stats

Heterozygosity graphs show the expected and observed heterozygosities (Hexp and Hobs, respectively) for each locus in the populations.

I used Bartlett test of homogeneity of variances to test if the expected and observed heterozygosities were similar.

In these tables, where it shows 'NaN' it means that the parameter cannot be calculated.

In the case of Mali, there is only one region and one population location, so the values calculated for the country, the region and the population are the same.

### Countries

```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
library(knitr)

tables_summary$countries
new_basicstats$countries
plots_ho_he$countries

cat("  \n\n### Types of land use\n\n")

tables_summary$use
new_basicstats$use
plots_ho_he$use

for (i in names(tables_summary)[3:5]) {
	
	if (i == "MALI") {next}
	cat("  \n\n###", i, "  \n\n#### Regions \n\n  ")
	
	print(tables_summary[[i]]$regions)
	cat("\n")
	print(new_basicstats[[i]]$regions)
	cat("\n")
	print(plots_ho_he[[i]]$regions)
	
	
	cat("  \n\n#### 'Localities'\n\n  ")
	print(tables_summary[[i]]$pops)
	cat("\n")
	print(new_basicstats[[i]]$pops)
	cat("\n")
	print(plots_ho_he[[i]]$pops)
}

```

## Allelic richness

```{r, warning=F, echo=F}
load("RData/allelic_richness.RData")
population_information <- read.csv(
	file = "input data - DO NOT MODIFY/Population.csv")
names(population_information) <- c("Taxa", "Taxa.Name", "Country",
																	 "Region", "Location.Population",
																	 "Type.Land.Use")

population_information = 
	population_information[- which(is.na(population_information$Taxa)),]

sample_size = lapply(3:6, function(x) {
	a = table(population_information[,x])
	
	if(x %in% c(3, 6)) {
		names = sort(unique(population_information[, x]))
	} else if (x == 4) {
		names = paste0(population_information[,3],"_",
									 population_information[,4]) %>% unique %>% sort
	} else {
		names = paste0(population_information[,3],"_",
									 population_information[,4],"_",
									 population_information[,5]) %>% unique %>% sort	
	}
	
	data.frame(names, n = a)
})

names(sample_size) = names(longer_allelic)
longer_allelic$countries +
	geom_text(data = sample_size[[1]], aes(x = names, y = 1.75, 
																				 label = paste0("n = ", n.Freq)))+
	labs(title = "Allelic richness for Countries", x = "") +
	theme(axis.text.x = element_text(vjust = 0.5, hjust = 1))

longer_allelic$regions +
	labs(title = "Allelic richness for Regions", x = "", 
			 caption = "Value inside indicates sample size")+
	geom_text(data = sample_size[[2]], aes(x = names, y = 1.75, 
																				 label = paste0(n.Freq)))+
	theme(axis.text.x = element_text(vjust = 0.5, hjust = 1))

longer_allelic$populations +
	geom_text(data = sample_size[[3]], aes(x = names, y = 1.75, 
																				 label = paste0(n.Freq)))+
	labs(title = "Allelic richness for Population localities", x = "Locality", 
			 caption = "Value inside indicates sample size")+
	theme(axis.text.x = element_text(vjust = 0.5, hjust = 1))

longer_allelic$use +
	geom_text(data = sample_size[[4]], aes(x = names, y = 1.75, 
																				 label = paste0("n = ", n.Freq)))+
	labs(title = "Allelic richness for Land uses", x = "Type of land use")+
	theme(axis.text.x = element_text(vjust = 0.5, hjust = 1))

```

# Genetic distance between samples

## PCA

I used Tassel5 to perfom principal component analysis on the SNP data. Each component explains very little of the variance of the data, so here are two graphs with the first 4 principal components.

```{r echo=FALSE, message=FALSE, warning=FALSE}
theme_set(theme_minimal() + 
						theme(axis.line = element_blank()))

data.frame(eigenvectors = 1:length(karite_ev$names.arg),
					 variance_explained = karite_ev$names.arg) %>%
	ggplot(aes(eigenvectors, variance_explained)) +
	geom_col(fill = "black", color = "black") +
	scale_y_continuous(breaks = seq(0, 10, 1)) +
	labs(x = "Eigenvectors", y = karite_ev[[2]])

cat("    \n\n")
karite_pca + 
	geom_hline(yintercept = c(-5, 5), color = "red", linetype = "dotdash") +
	geom_vline(xintercept = c(-9), color = "red", linetype = "dotdash")
	
```

I repeated the PCA with only the individuals in the big group (those located between the red dotdashed line in PC2 & to the right of the red dotdashed line in PC1).

```{r echo=FALSE, message=FALSE, warning=FALSE}

data.frame(eigenvectors = 1:length(biggroup_ev$names.arg),
					 variance_explained = biggroup_ev$names.arg) %>%
	ggplot(aes(eigenvectors, variance_explained)) +
	geom_col(fill = "black", color = "black") +
	scale_y_continuous(breaks = seq(0, 10, 0.5)) +
	labs(x = "Eigenvectors", y = karite_ev[[2]])

cat("    \n\n")
biggroup_pca_plot

```

## Correspondance analysis

```{r Correspondence Analysis, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, results="asis"}
theme_set(theme_minimal() + 
						theme(axis.line = element_blank()))

data.frame(pops = rownames(coa_pops[[3]]$li), 
													coa_pops[[3]]$li) %>%
		mutate(Country = gsub("([A-Z]+)_.+", "\\1", pops)) %>%
		ggplot(aes(Axis1, Axis2, fill = Country, label = pops)) +
		geom_label() +
		labs(x = paste0("Axis 1: (", round(coa_pops[[3]]$eig[1]*100, 2), "%)"),
				 y = paste0("Axis 2: (", round(coa_pops[[3]]$eig[2]*100, 2), "%)"),
				 title = "Correspondence Analysis") +
		geom_hline(yintercept = 0) +
		geom_vline(xintercept = 0)

```

## AMOVA

```{r AMOVA, echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, results="asis"}
load("RData/amova_res.RData")
knitr::kable(
	amova$results, 
	caption = 
		"Result AMOVA. Hierarchy: _Country - Region - Location.Population_.")
knitr::kable(amova$componentsofcovariance, 
						 caption = "Components of covariance.")
knitr::kable(amova$statphi, 
						 caption = "Phi - population variation statistics.")

data.frame(Test = rand_amova$names,
					 Obs = rand_amova$obs, 
					 Std.Obs = rand_amova$expvar$Std.Obs,
					 Alter = rand_amova$alter,
					 p.Value = rand_amova$pvalue) %>%
	knitr::kable(caption = "Monte-Carlo tests (999 repetitions). Sample = population localities")

```
## Nei's distance 

I used the function "dist.genpop" from the package *adegenet* to calculate the Nei's distance between the samples.

In the heatmap below, darkgray to black indicates low genetic distance between the populations, while light gray to white indicates high genetic distance.

We can see they are very similar but there are some groups:

-   The samples from Tchad: Moyen-Chari - Kagala, Mayo-kebbi - Pala, Mandoul - Peni and Mandoul - Kemkiam (top right)
-   The sample from Mali with samples from Cameroun Ouest (center)
-   The samples from Cameroun Nord (down and left from the previous ones)
-   Two samples from Cameroun Extreme-Nord: Mayo-Tsana and Mayo-Dana (down and left from the previous ones)

We also see that the samples from Tchad Logone Oriental are different from all the others and from each other.

```{r echo=FALSE, fig.height=8, fig.width=9, message=FALSE, warning=FALSE}
load("RData/heatmap_nei.RData")

heatmap_nei
```

## Differentiation metrics

I used the package _mmod_ to calculate the Jost's D, Hendrick's G'st and Nei's G'st between all populations.

### Countries

```{r echo=FALSE, fig.height=6, message=F, warning=F, results="asis"}
library(ggpubr)

for (i in names(differentiation)) {
	a = ifelse(i == "Jost_D", "Jost's D",
						 ifelse(i == "Nei_Gst", "Nei's G'st", "Hendrick's G'st"))

	knitr::kable(differentiation[[i]]$countries[
	, -ncol(differentiation[[i]]$countries)],
	caption = a) %>% print
}

# ggarrange(plotlist = plots_longer[[1]], 
# 					ncol = 2, nrow = 2, align = "h")


```


```{r echo=FALSE, message=F, warning=F, results="asis"}
# for (i in names(differentiation)) {
# 	a = ifelse(i == "Jost_D", "Jost's D", 
# 						 ifelse(i == "Nei_Gst", "Nei's G'st", "Hendrick's G'st"))
# 	
# 	# knitr::kable(differentiation[[i]]$regions[
# 	# , -ncol(differentiation[[i]]$regions)], 
# 	# caption = a) %>% print
# }
# 
# plots_longer[[2]][[1]]
# plots_longer[[2]][[2]]
# plots_longer[[2]][[3]]

```


```{r echo=FALSE, fig.height=6, fig.width=8, message=FALSE, warning=FALSE, results="asis"}
library(xlsx)

for (i in names(differentiation)) {
	a = ifelse(i == "Jost_D", "Jost's D",
						 ifelse(i == "Nei_Gst", "Nei's G'st", "Hendrick's G'st"))
	write.xlsx(differentiation[[i]]$populations, 
						 file = "Differentiation_metrics_population.xlsx", 
						 sheetName=a, 
						 col.names=TRUE, row.names=TRUE, append=FALSE)
}

# plots_longer[[3]][[1]]
# plots_longer[[3]][[2]]
# plots_longer[[3]][[3]]
# 
```

### Type of land use

```{r echo=FALSE, fig.height=3, fig.width=5, message=FALSE, warning=FALSE, results="asis"}
for (i in names(differentiation)) {
	a = ifelse(i == "Jost_D", "Jost's D", 
						 ifelse(i == "Nei_Gst", "Nei's G'st", "Hendrick's G'st"))
	
	knitr::kable(differentiation[[i]]$use[
	, -ncol(differentiation[[i]]$use)], 
	caption = a, digits = 5) %>% print
}

# plots_longer[[4]][[1]]
# cat("    \n\n")
# plots_longer[[4]][[2]]
# cat("    \n\n")
# 
# plots_longer[[4]][[3]]

```

## UPGMA 

```{r UPGMA, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
library(ape)
use_pack = c("phylogram", "pvclust")
for (i in use_pack) {
	#install.packages(i, dependencies = T); 
	library(i, character.only = T)}

load(file = "RData/tree_karite.RData")
#main = "Genetic distance (proportion of loci that are different)" 
library(dendextend)
colors = ifelse(grepl("TCHAD", tree_karite$tip.label), 
								"red", ifelse(grepl("CAM", tree_karite$tip.label), "black",
															"blue"))
plot(tree_karite, show.node.label = T, cex = 0.8, tip.color = colors,font = 1, 
		 main = "UPGMA tree for population localities")
mtext(side=1, line=1, adj=0, cex=0.7, "Values indicate node support using 500 samples.")
mtext(side=1, line=2, adj=0, cex=0.7, "Calculated using Euclidean distance matrix")
```

